---
title: "exploratory analysis of auto-regression of biosensors"
output: html_document
date: "2023-11-10"
---


# Overview
dysregulation in physio system is a predictor of aggressive behavior (or sib, ed)
dysregulation is measured by dynamics in EDA, hand movements, bvp

To examine: effect of AR(1) of EDA, accelerometer, BVP on aggressive behavior
Expect: Higher AR indicates emotion dysregulation, and elicits higher aggressive behavior

Prelim result:
Precision-recall is low for prediction of aggressive behavior based on biosensor AR, including xgboosting, multilevel modeling (generalized)

# accelerometr reference
https://www.google.com/search?sca_esv=589973973&rlz=1C1ONGR_enUS1027US1027&sxsrf=AM9HkKkRsOOsMIKoQZlYlspnVnxPyMNxhQ:1702343399510&q=accelerometer+xyz&tbm=isch&source=lnms&sa=X&ved=2ahUKEwj32POs24iDAxUqFVkFHc2EBkEQ0pQJegQICxAB&biw=1360&bih=643&dpr=1


````{r}

library(ggplot2)
library(dplyr)
library(lmerTest)
library(lme4)
library(lubridate)
library(psych)
library(pompom)

library(zoo)
library(xgboost)
library(reshape2)

library(pROC)
library(PRROC)

library(lme4)
library(lmerTest)

library(caret)
set.seed(1234)
```
 


```{r}
path <- getwd()
setwd(path)
# all_data <- read.csv("all_data.csv", nrows = 200)
all_data <- read.csv("all_data.csv")

all_data$episode <- as.numeric(substring(all_data$file,9,10))

```

# check distribution of data
```{r}
ggplot(data = all_data,aes(x = eda))+
  geom_histogram()

all_data$eda_zero <- ifelse(abs(all_data$eda) <0.01, 1,0)
table(all_data$eda_zero) 

all_data$bvp_zero <- ifelse(abs(all_data$bvp)<0.01, 1,0)
table(all_data$bvp_zero) 

table(all_data$agg) # 0.5% is 1
table(all_data$ed) # 4.5% is 1
table(all_data$sib) # 1.8% is 1

```

# Compute mean level of 3 behaviors (agg, sib, ed) as well as auto-regression (AR) coefficient of biosensors (EDA, BVP, ACC)



```{r}

id_list <- unique(all_data$id)
time_window <- 180 # seconds
predict_window <- 60 # seconds
sampling_rate <- 4
sliding_unit <- 15 * sampling_rate # data points, 60 seconds

window_data_collect <- NULL

for (id in id_list){
  id_data <- all_data[all_data$id == id,]
  session_list <- unique(id_data$session)
  for (session in session_list){
    id_session_data <- id_data[id_data$session == session,]
    episode_list <- unique(id_session_data$episode)
  
    for (episode in episode_list){
      id_episode_data <- id_session_data[id_session_data$episode == episode,]
      id_episode_data$timestamp2 <-as.POSIXct(id_episode_data$timestamp/1000, 
                                   origin="1970-01-01",
                                   tz = "UTC")
      id_episode_data$timestamp3 <-as.POSIXct(id_episode_data$timestamp2,
                                   format = "%Y-%m-%d %H:%M:%OS")
        
      id_episode_data <- id_episode_data[order(id_episode_data$timestamp3),]
      # id_episode_data$eda_diff <- id_episode_data$eda - lag(id_episode_data$eda,1)
      
      # compute distance from step to step
      id_episode_data$acc_distance <- sqrt(
        (id_episode_data$acc_x - lag(id_episode_data$acc_x,1))^2 + 
          (id_episode_data$acc_y - lag(id_episode_data$acc_y,1))^2 + 
          (id_episode_data$acc_z - lag(id_episode_data$acc_z,1))^2)
      
  
      
      if (time_window*sampling_rate+1  < 
          nrow(id_episode_data) -
          predict_window*sampling_rate  - 
          sliding_unit){
        for (row in seq(time_window*sampling_rate+1 ,
                        nrow(id_episode_data) - 
                        predict_window*sampling_rate , 
                        sliding_unit)
             ){
            
          # select corresponding rows for the 3 minutes
          usem_data <- id_episode_data[(row
                                  -time_window*sampling_rate)
                                 :row, ]
          # select column
          usem_data <- usem_data[c("eda","bvp","acc_distance")]
          # remove na
          # usem_data <- usem_data[complete.cases(usem_data),]
          
          mean_eda <- mean(usem_data$eda, na.rm = T)
          mean_acc_distance <- mean(usem_data$acc_distance, na.rm = T)
          mean_bvp <- mean(usem_data$bvp, na.rm = T)
          
          eda_ar <- coefficients(lm(eda~lag(eda,1),data = usem_data))[2]
          acc_ar <- coefficients(lm(acc_distance~lag(acc_distance,1),data = usem_data))[2]
          bvp_ar <- coefficients(lm(bvp~lag(bvp,1),data = usem_data))[2]
          
     
            predict_agg<- mean(id_episode_data$agg
                               [row: (row +predict_window*sampling_rate)], 
                               na.rm = T)
            
            predict_sib<- mean(id_episode_data$sib
                               [row: (row +predict_window*sampling_rate)], 
                               na.rm = T)
            predict_ed<- mean(id_episode_data$ed
                              [row: (row +predict_window*sampling_rate)], 
                              na.rm = T)
           
           
          window_data_collect <- rbind(window_data_collect, 
                                       c(id, session, episode,row,
                                         predict_agg,
                                         predict_sib,
                                         predict_ed, 
                                         mean_eda, 
                                         mean_acc_distance,
                                         mean_bvp,
                                         eda_ar,
                                         acc_ar,
                                         bvp_ar
                                         ))
        }# end of window sliding loop
        
        
      }else{
      next
      }# end of window size exceeding rows condition
        
      }# end of  episode loop
    } # end of session loop
} # end of id loop


window_data_collect <- data.frame(window_data_collect)
names(window_data_collect) <-c ("id", "session", "episode","row",
                               
                               "predict_agg",
                               "predict_sib",
                               "predict_ed",
                        
                               "mean_eda", 
                               "mean_acc_distance",
                               "mean_bvp",
                                "eda_ar",
                                 "acc_ar",
                                 "bvp_ar")

write.csv(window_data_collect,"exploratory_ar_df.csv" )


```






